---
title: "Predicting House Prices using MLR"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center', out.width = '80%')
library(tidyverse)
library(car)
library(leaps)
library(corrplot)
library(ggplot2)
library(gridExtra)
library(moments)
library(dplyr)
library(patchwork)
library(ggcorrplot)
library(MASS)
library(leaps)
```

# Data Cleaning

```{r}
df <- read.csv("HousePriceTrain.csv", stringsAsFactors = FALSE)
```

## Full Model before Cleaning

We have to remove Utilities first because it's a factor with only one level.

```{r}
# clean up columns with only one value
constant_cols <- sapply(df, function(x) length(unique(x)) <= 1)
removed_names <- names(df)[constant_cols]
print(paste("Removing constant columns:", paste(removed_names, collapse = ", ")))
df <- df[, !constant_cols]
```
```{r}
uncleaned_full_model <- lm(SalePrice ~ . - X - House_Review_Text, data = df)
summary(uncleaned_full_model)
par(mfrow = c(2, 2))
plot(uncleaned_full_model)

selected_vars_for_corr <- df %>% 
  dplyr::select(SalePrice, GrLivArea, TotalBsmtSF, GarageArea, 
         OverallQual, OverallCond, House_Age)
pairs(selected_vars_for_corr)
```

```{r}
clean_house_data <- function(df, is_training = TRUE) {
  
  # 1. Temporal Processing
  df$YrSold <- round(df$YrSold)
  df$YearRemodAdd <- round(df$YearRemodAdd)
  df$GarageYrBlt <- round(df$GarageYrBlt)
  
  df$Remodel_Age <- pmax(0, df$YrSold - df$YearRemodAdd)
  df$GaragePresent <- as.integer(df$Garage == "Yes")
  df$Garage_Age <- ifelse(df$GaragePresent == 1, pmax(0, df$YrSold - df$GarageYrBlt), 0)
  
  # 2. Spatial & Categorical
  spatial_vars <- c("Neighborhood", "Condition1", "Condition2", "LotConfig", "LandContour", "LandSlope")
  df[spatial_vars] <- lapply(df[spatial_vars], factor)
  
  # 3. Count-Based Variables
  count_vars <- c("BsmtFullBath", "BsmtHalfBath", "FullBath", "HalfBath", 
                  "BedroomAbvGr", "KitchenAbvGr", "TotRmsAbvGrd", "Fireplaces", "GarageCars")
  for (col in count_vars) { df[[col]] <- as.numeric(round(df[[col]])) }
  
  categorical_vars <- c("MSZoning", "Street", "Alley", "LandContour", "LotConfig", 
                        "Neighborhood", "Condition1", "Condition2", "BldgType", "HouseStyle", 
                        "RoofStyle", "RoofMatl", "Exterior1st", "Exterior2nd", "MasVnrType", 
                        "Foundation", "Heating", "Electrical", "Functional", "GarageType", 
                        "PavedDrive", "MiscFeature", "SaleType", "SaleCondition")
  
  df$MSSubClass <- as.factor(as.character(round(df$MSSubClass)))
  for (col in categorical_vars) { df[[col]] <- as.factor(as.character(df[[col]])) }
  
  # 4. Ordinal Variables
  
  # Ordinal variables are categorical variables with a natural order but unknown spacing between levels. We identified ordinal predictors by examining unique
  # levels and assessing whether the categories represent increasing or decreasing quality, condition, or intensity.
  
  # Rounded OverallQual and OverallCond back to their original 1–10 ordinal scale and bounded the values between 1 and 10. The raw data contained small decimal 
  # noise, so this step restores their intended discrete rating structure before modeling.
  
  # Converting all ordinal categorical variables into ordered factors and explicitly defined their natural ranking to make sure the model recognizes the correct 
  # order of these categories (e.g., Poor < Fair < Typical < Good < Excellent) instead of treating them as nominal variables.
  
  df$OverallQual <- pmin(pmax(round(df$OverallQual), 1), 10)
  df$OverallCond <- pmin(pmax(round(df$OverallCond), 1), 10)
  
  lvl_qual_5   <- c("Po","Fa","TA","Gd","Ex")                  
  lvl_func     <- c("Sev","Maj2","Maj1","Mod","Min2","Min1","Typ")
  lvl_gfinish  <- c("Unf","RFn","Fin")
  lvl_paved    <- c("N","P","Y")
  lvl_lotshape <- c("IR3","IR2","IR1","Reg")
  lvl_landslp  <- c("Sev","Mod","Gtl")
  lvl_bexp     <- c("No","Mn","Av","Gd")
  lvl_bfin     <- c("Unf","LwQ","Rec","BLQ","ALQ","GLQ")
  lvl_fence    <- c("MnWw","GdWo","MnPrv","GdPrv")             
  lvl_pool     <- c("Fa","Gd","Ex")
  
  # Applying factors (using a list to avoid repetition)
  df$ExterQual <- factor(df$ExterQual, levels = lvl_qual_5, ordered = TRUE)
  df$ExterCond <- factor(df$ExterCond, levels = lvl_qual_5, ordered = TRUE)
  df$BsmtQual  <- factor(df$BsmtQual,  levels = lvl_qual_5, ordered = TRUE)
  df$BsmtCond  <- factor(df$BsmtCond,  levels = lvl_qual_5, ordered = TRUE)
  df$HeatingQC <- factor(df$HeatingQC, levels = lvl_qual_5, ordered = TRUE)
  df$KitchenQual <- factor(df$KitchenQual, levels = lvl_qual_5, ordered = TRUE)
  df$FireplaceQu <- factor(df$FireplaceQu, levels = lvl_qual_5, ordered = TRUE)
  df$GarageQual <- factor(df$GarageQual, levels = lvl_qual_5, ordered = TRUE)
  df$GarageCond <- factor(df$GarageCond, levels = lvl_qual_5, ordered = TRUE)
  df$Functional <- factor(df$Functional, levels = lvl_func, ordered = TRUE)
  df$GarageFinish <- factor(df$GarageFinish, levels = lvl_gfinish, ordered = TRUE)
  df$PavedDrive <- factor(df$PavedDrive, levels = lvl_paved, ordered = TRUE)
  df$LotShape <- factor(df$LotShape, levels = lvl_lotshape, ordered = TRUE)
  df$LandSlope <- factor(df$LandSlope, levels = lvl_landslp, ordered = TRUE)
  df$BsmtExposure <- factor(df$BsmtExposure, levels = lvl_bexp, ordered = TRUE)
  df$BsmtFinType1 <- factor(df$BsmtFinType1, levels = lvl_bfin, ordered = TRUE)
  df$BsmtFinType2 <- factor(df$BsmtFinType2, levels = lvl_bfin, ordered = TRUE)
  df$Fence <- factor(df$Fence, levels = lvl_fence, ordered = TRUE)
  df$PoolQC <- factor(df$PoolQC, levels = lvl_pool, ordered = TRUE)
  
  # 5. Binary
  
  # Converted all confirmed binary categorical variables into numeric 1/0 indicators for consistent model input and easier interpretation.
  
  df$CentralAir <- ifelse(df$CentralAir == "Y", 1, 0)
  df$Central_AC_Heater <- ifelse(df$Central_AC_Heater == "Yes", 1, 0)
  df$Basement <- ifelse(df$Basement == "Yes", 1, 0)
  df$Garage <- ifelse(df$Garage == "Yes", 1, 0)
  df$Street <- ifelse(df$Street == "Pave", 1, 0)
  df$Alley <- ifelse(df$Alley == "Pave", 1, 0)
  
  # 6. NLP (Sentiment)
  pos_words <- c("excellent", "upgrades", "lovely", "quiet", "beautiful", "well", "maintained", "friendly", "spacious", "modern", "great", "renovated", "luxury", "new")
  neg_words <- c("issues", "repairs", "poor", "old", "outdated", "worn", "noisy", "traffic", "crowded", "small", "dated", "needs", "limited")
  
  get_sentiment <- function(text) {
    if (is.na(text) || text == "") return(0)
    text <- str_to_lower(text)
    pos_score <- str_count(text, paste0("\\b(", paste(pos_words, collapse = "|"), ")\\b"))
    neg_score <- str_count(text, paste0("\\b(", paste(neg_words, collapse = "|"), ")\\b"))
    return(pos_score - neg_score)
  }
  
  df$ReviewSentiment <- factor(sapply(df$House_Review_Text, get_sentiment))
  
  # 7. Feature Engineering
  df$TotalArea <- df$GrLivArea + df$TotalBsmtSF + df$GarageArea + df$WoodDeckSF + df$OpenPorchSF + df$EnclosedPorch + 
                  df$X3SsnPorch + df$ScreenPorch + df$PoolArea
  df$OutdoorArea <- df$WoodDeckSF + df$OpenPorchSF + df$EnclosedPorch + df$X3SsnPorch + df$ScreenPorch + df$PoolArea
  
  df$AboveBath <- df$FullBath + 0.5*df$HalfBath
  df$BsmtBath <- df$BsmtFullBath + 0.5*df$BsmtHalfBath
  
  df$HasFireplace <- as.integer(df$Fireplaces > 0)
  df$Fireplaces_cap2 <- pmin(df$Fireplaces, 2)
  
  df$House_Age <- pmax(0, df$House_Age)
  
  df$GarageCars_cap4 <- pmin(df$GarageCars, 4)
  
  df$Kitchen2plus <- as.integer(df$KitchenAbvGr >= 2)
  
  # 8. Interactions
  
  # Because there are floating point issues, we reconstruct the two variables using our rounded data.
  
  df$Qual_x_Area <- df$OverallQual * df$GrLivArea
  df$Garage_x_Qual <- df$OverallQual * df$GarageCars
  df$Qual_x_TotalArea <- df$OverallQual * df$TotalArea
  df$Cond_x_TotalArea <- df$OverallCond * df$TotalArea
  df$House_Age_x_TotalArea <- log1p(df$House_Age) * df$TotalArea
  
  # 9. Final Cleanup
  
  # We drop the terms that were used for feature engineering and interaction terms and doesn't say much on themselves.
  
  cols_to_remove <- c("WoodDeckSF", "OpenPorchSF", "EnclosedPorch", "X3SsnPorch", "ScreenPorch", "PoolArea", 
                      "YrSold", "YearBuilt", "YearRemodAdd", "ExterCond", "GarageCond", "BsmtCond", 
                      "X1stFlrSF", "X2ndFlrSF", "GarageYrBlt", "MoSold", "Basement", "Garage", 
                      "TotalBsmtSF", "LowQualFinSF", "GaragePresent", "House_Review_Text", "BsmtFullBath",
                      "BsmtHalfBath", "FullBath", "HalfBath", "Fireplaces", "GarageCars", "KitchenAbvGr")
  
  df_clean <- df %>% dplyr::select(-all_of(cols_to_remove))
  
  return(df_clean)
}
```

```{r}
df_clean <- clean_house_data(df)
write.csv(clean_house_data(df), "HousePriceTrain_Cleaned.csv", row.names = FALSE)
```

---

# Data Description

## Response Variable EDA

```{r}
summary(df_clean$SalePrice)
```

We examine the distribution of `SalePrice` because MLR assumes normality.

```{r}
library(ggplot2)
library(gridExtra)

# Plot 1: Original Distribution
p1 <- ggplot(df_clean, aes(x = SalePrice)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of SalePrice", x = "Sale Price", y = "Frequency") +
  scale_x_continuous(labels = scales::comma)

# Plot 2: Log Transformation
p2 <- ggplot(df_clean, aes(x = log(SalePrice))) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Log(SalePrice)", x = "Log(Sale Price)", y = "Frequency")

grid.arrange(p1, p2, ncol = 2)

original_skew <- moments::skewness(df_clean$SalePrice, na.rm = TRUE)
log_skew <- moments::skewness(log(df_clean$SalePrice), na.rm = TRUE)

cat("Original Skewness:", original_skew, "\n")
cat("Log-Transformed Skewness:", log_skew, "\n")
```

`SalePrice` is highly right skewed, suggesting that a log transformation is needed.


```{r}
df_clean$Log_SalePrice <- log(df_clean$SalePrice)
summary(df_clean$Log_SalePrice)
```

## Preliminary Predictor Selection Logic

Since there are about 90 variables in the data set, we will focus mainly on the follow key variables during EDA based on typical housing economics:

- `GrLivArea`, `TotalBsmtSF`, `GarageArea`
- `OverallQual`, `OverallCond`
- `Neighborhood`
- `HouseAge` (constructed with `YrSold` and `YearBuilt`)

## Correlation Analysis (Numerical Predictors)

We will use a correlation matrix to identify which numerical variables have the strongest linear relationship with `Log_SalePrice`.

```{r}
numeric_df <- df_clean %>% dplyr::select(where(is.numeric))

correlations <- cor(numeric_df, use = "complete.obs")
log_price_correlations <- sort(correlations[,"Log_SalePrice"], decreasing = TRUE)

# Top 10 most influential numeric variables for Log_SalePrice
top_numeric <- log_price_correlations[!names(log_price_correlations) %in% c("Log_SalePrice", "SalePrice")]
head(top_numeric, 10)
```

```{r}
major_vars <- c("GrLivArea", "BsmtFinSF1", "BsmtFinSF2", "GarageArea", 
                "OverallQual", "OverallCond", "Neighborhood", "House_Age")

major_numeric <- c("GrLivArea", "BsmtFinSF1", "BsmtFinSF2", "GarageArea", 
                   "OverallQual", "OverallCond", "House_Age")

cor_matrix <- cor(df_clean[major_numeric], df_clean$Log_SalePrice, use = "complete.obs")
colnames(cor_matrix) <- "Log_SalePrice"
print(cor_matrix)

```
```{r}
p_living <- ggplot(df_clean, aes(x = GrLivArea, y = Log_SalePrice)) +
  geom_point(alpha = 0.1, color = "steelblue") +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Living Area", x = "Sq Ft", y = "Log Price") +
  theme_minimal()

p_bsmt1 <- ggplot(df_clean, aes(x = BsmtFinSF1, y = Log_SalePrice)) +
  geom_point(alpha = 0.1, color = "steelblue") +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Basement Area 1", x = "Sq Ft", y = "") +
  theme_minimal()

p_bsmt2 <- ggplot(df_clean, aes(x = BsmtFinSF2, y = Log_SalePrice)) +
  geom_point(alpha = 0.1, color = "steelblue") +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Basement Area 2", x = "Sq Ft", y = "") +
  theme_minimal()

p_garage <- ggplot(df_clean, aes(x = GarageArea, y = Log_SalePrice)) +
  geom_point(alpha = 0.1, color = "steelblue") +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Garage Area", x = "Sq Ft", y = "") +
  theme_minimal()


p_living + p_bsmt1 + p_bsmt2 + p_garage
```

```{r}
# for visualization, round them to 1-10
p_qual <- ggplot(df_clean, aes(x = as.factor(round(OverallQual)), y = Log_SalePrice)) +
  geom_boxplot(fill = "steelblue", outlier.alpha = 0.2) +
  labs(title = "Overall Quality (Rounded as 1-10)", 
       x = "Quality Rating (1-10)", 
       y = "Log_SalePrice") +
  theme_minimal()

# 2. Boxplot for Overall Condition
p_cond <- ggplot(df_clean, aes(x = as.factor(round(OverallCond)), y = Log_SalePrice)) +
  geom_boxplot(fill = "steelblue", outlier.alpha = 0.2) +
  labs(title = "Overall Condition (Rounded as 1-10)", 
       x = "Condition Rating (1-10)", 
       y = "Log_SalePrice") +
  theme_minimal()

# Age Scatter plot
p_age <- ggplot(df_clean, aes(x = House_Age, y = Log_SalePrice)) +
  geom_point(alpha = 0.1, color = "steelblue") +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "House Age", x = "Years", y = "") + 
  theme_minimal()

# Combine plots
p_qual / (p_cond + p_age)
```

```{r}
df_clean %>%
  mutate(Neighborhood = reorder(Neighborhood, Log_SalePrice, FUN = median)) %>%
  ggplot(aes(x = Neighborhood, y = Log_SalePrice)) +
  geom_boxplot(fill = "steelblue") +
  coord_flip() +
  labs(title = "Neighborhood Impact on Log_SalePrice",
       subtitle = "Ordered by Median Log Price",
       x = "Neighborhood", y = "Log_SalePrice") +
  theme_minimal()
```

```{r}
vars_for_corr <- df_clean %>% 
  dplyr::select(Log_SalePrice, GrLivArea, BsmtFinSF1, GarageArea, 
         OverallQual, OverallCond, House_Age)
corr_matrix <- cor(vars_for_corr, use = "complete.obs")
ggcorrplot(corr_matrix, 
           hc.order = TRUE, 
           type = "lower",
           lab = TRUE, 
           title = "Correlation Heatmap of Key Predictors",
           colors = c("#6D9EC1", "white", "#E46726"))
```

---
```{r}
# Improved Model: Balancing interpretability with predictive power
improved_model <- lm(log(SalePrice) ~ 
                       # --- Transformed Continuous & Engineered ---
                       I(Qual_x_TotalArea^(1/3)) + 
                       log(GrLivArea) + 
                       log(LotArea) + 
                       I(House_Age^(1/3)) +
                       GarageArea +
                       log1p(BsmtFinSF1) +
                       
                       # --- Quality & Condition (Ordinal) ---
                       OverallQual + 
                       OverallCond + 
                       BsmtQual + 
                       KitchenQual + 
                       ExterQual +
                       
                       # --- Categorical Blocks (Location & Type) ---
                       Neighborhood +   # Crucial for location
                       MSZoning +       # Land use type
                       BldgType +       # House structure type
                       Foundation +     # Structural quality
                       SaleCondition +  # Market context
                       Functional,      # Home functionality rating
                     data = df_clean)


summary(improved_model)
par(mfrow = c(2, 2))
plot(improved_model)
vif(improved_model)

```

```{r}
vars_model <- c(
"SalePrice",
"Qual_x_TotalArea",
"GrLivArea",
"LotArea",
"House_Age",
"GarageArea",
"BsmtFinSF1",
"OverallQual",
"OverallCond",
"BsmtQual",
"KitchenQual",
"ExterQual",
"Neighborhood",
"MSZoning",
"BldgType",
"Foundation",
"SaleCondition",
"Functional"
)

df_model <- df_clean[, vars_model]

num_vars <- df_model[, sapply(df_model, is.numeric), drop = FALSE]

```

```{r}
# Correlation Matrix of Variables in Initial Model
pairs(num_vars,
      col = "darkblue",
      pch = 16,
      cex = 0.3,
      main = "Correlation Matrix of Variables in Initial Model")
```

```{r}
#  Distribution of Continuous Variables
library(ggplot2)
library(tidyr)

cont_vars <- c(
"SalePrice",
"Qual_x_TotalArea",
"GrLivArea",
"LotArea",
"House_Age",
"GarageArea",
"BsmtFinSF1"
)

df_long <- df_clean[, cont_vars] |>
  pivot_longer(cols = everything(),
               names_to = "Variable",
               values_to = "Value")

ggplot(df_long, aes(x = Value)) +
  geom_histogram(bins = 30,
                 fill = "steelblue",
                 color = "white") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal()

```


```{r}
# Distribution of Ordinal Variables
ord_vars <- c(
"OverallQual",
"OverallCond",
"BsmtQual",
"KitchenQual",
"ExterQual"
)

df_long_ord <- df_clean[, ord_vars] |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(cols = everything(),
               names_to = "Variable",
               values_to = "Value")

ggplot(df_long_ord, aes(x = Value)) +
  geom_bar(fill = "steelblue") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal()
```

```{r}
# Distribution of Categorical Variables
cat_vars <- c(
"Neighborhood",
"MSZoning",
"BldgType",
"Foundation",
"SaleCondition",
"Functional"
)

df_long_cat <- df_clean[, cat_vars] |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

ggplot(df_long_cat, aes(x = Value)) +
  geom_bar(fill = "steelblue") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


---

# Results and Interpretation

## Full Model

```{r}
options(contrasts = c("contr.treatment", "contr.treatment"))
```

```{r}
full_model <- lm(SalePrice ~ . - Log_SalePrice - X -  Log_GrLivArea - Log_LotArea, data = df_clean)
boxCox(full_model)
```

```{r}
full_model <- lm(Log_SalePrice ~ . - SalePrice - X - Log_GrLivArea - Log_LotArea, data = df_clean)
formula(full_model)
```

## Power Transform

```{r}
df_numeric <- df_clean %>% dplyr::select(where(is.numeric))

safe_vars <- c(
  "LotFrontage", 
  "LotArea", 
  "GrLivArea", 
  "BsmtFinSF1", 
  "BsmtUnfSF", 
  "GarageArea", 
  "SalePrice",
  "House_Age",
  "OutdoorArea",
  "TotalArea",
  "MasVnrArea",
  "BsmtFinSF2",
  "Remodel_Age"
)

pt_final <- powerTransform(as.matrix(df_numeric[, safe_vars]) ~ 1, family = "yjPower")
summary(pt_final)
```


```{r}
df_clean <- df_clean %>% dplyr::select(-SalePrice)

transformed_model <- lm(Log_SalePrice ~ . + log1p(LotArea) + log1p(LotFrontage) + log1p(BsmtFinSF1) + log1p(BsmtUnfSF) + log1p(GarageArea) + log1p(House_Age) + log1p(OutdoorArea) + log1p(TotalArea) + log1p(MasVnrArea) + log1p(BsmtFinSF2) + log1p(Remodel_Age) - X - GrLivArea - LotArea - Log_GrLivArea - Log_LotArea - LotFrontage - BsmtFinSF1 - BsmtUnfSF - GarageArea - OutdoorArea - TotalArea - House_Age - MasVnrArea - BsmtFinSF2 - Remodel_Age, data = df_clean)

formula(transformed_model)
```

```{r}
p <- ncol(model.matrix(transformed_model)) - 1
p
```

```{r}
bad <- names(coef(transformed_model))[is.na(coef(transformed_model))]
bad
```

```{r}
transformed_model_new <- update(transformed_model, . ~ . - BldgType)
names(coef(transformed_model_new))[is.na(coef(transformed_model_new))]
v <- vif(transformed_model_new)
v[v[, "GVIF^(1/(2*Df))"] > 5, , drop = FALSE]
```

```{r}
transformed_model_reduced <- update(transformed_model_new, . ~ . - Garage_x_Qual - Qual_x_TotalArea - Cond_x_TotalArea - House_Age_x_TotalArea)
v2 <- vif(transformed_model_reduced)
v2[v2[, "GVIF^(1/(2*Df))"] > 5, , drop = FALSE]
```

```{r}
par(mfrow = c(2, 2))
plot(transformed_model_reduced)
```

```{r}
StanRes1 <- rstandard(transformed_model_reduced)
mf <- model.frame(transformed_model_reduced)

par(mfrow = c(2, 3))

plot(mf[["log1p(LotArea)"]], StanRes1,
     xlab = "log1p(LotArea)", ylab = "Standardized Residuals",
     main = "log1p(LotArea)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(LotFrontage)"]], StanRes1,
     xlab = "log1p(LotFrontage)", ylab = "Standardized Residuals",
     main = "log1p(LotFrontage)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(BsmtFinSF1)"]], StanRes1,
     xlab = "log1p(BsmtFinSF1)", ylab = "Standardized Residuals",
     main = "log1p(BsmtFinSF1)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(BsmtUnfSF)"]], StanRes1,
     xlab = "log1p(BsmtUnfSF)", ylab = "Standardized Residuals",
     main = "log1p(BsmtUnfSF)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(GarageArea)"]], StanRes1,
     xlab = "log1p(GarageArea)", ylab = "Standardized Residuals",
     main = "log1p(GarageArea)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(House_Age)"]], StanRes1,
     xlab = "log1p(House_Age)", ylab = "Standardized Residuals",
     main = "log1p(House_Age)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(OutdoorArea)"]], StanRes1,
     xlab = "log1p(OutdoorArea)", ylab = "Standardized Residuals",
     main = "log1p(OutdoorArea)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(TotalArea)"]], StanRes1,
     xlab = "log1p(TotalArea)", ylab = "Standardized Residuals",
     main = "log1p(TotalArea)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(MasVnrArea)"]], StanRes1,
     xlab = "log1p(MasVnrArea)", ylab = "Standardized Residuals",
     main = "log1p(MasVnrArea)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(BsmtFinSF2)"]], StanRes1,
     xlab = "log1p(BsmtFinSF2)", ylab = "Standardized Residuals",
     main = "log1p(BsmtFinSF2)")
abline(h = 0, lty = 2, col = 'red')

plot(mf[["log1p(Remodel_Age)"]], StanRes1,
     xlab = "log1p(Remodel_Age)", ylab = "Standardized Residuals",
     main = "log1p(Remodel_Age)")
abline(h = 0, lty = 2, col = 'red')
```

```{r}
avPlots(transformed_model_reduced, terms = ~ log1p(LotArea) + log1p(LotFrontage) + log1p(BsmtFinSF1) + log1p(BsmtUnfSF) + log1p(GarageArea) + log1p(House_Age) + log1p(OutdoorArea) + log1p(TotalArea) + log1p(MasVnrArea) + log1p(BsmtFinSF2) + log1p(Remodel_Age), layout = c(2, 3), ask = FALSE)
```

```{r}
n <- nrow(df_clean)
m0 <- lm(Log_SalePrice ~ 1, data = df_clean)
mfull <- transformed_model_reduced

step_aic <- step(
  m0,
  scope = list(lower = formula(m0), upper = formula(mfull)),
  direction = "both",
  k = 2,
  trace = 0
)

step_bic <- step(
  m0,
  scope = list(lower = formula(m0), upper = formula(mfull)),
  direction = "both",
  k = log(n),
  trace = 0
)
```

```{r}
formula(step_aic)
```

```{r}
formula(step_bic)
```

```{r}
term_count <- function(mod) {
  length(attr(terms(mod), "term.labels"))
}

trim_to_20_aic <- function(mod, max_terms = 20) {
  current <- mod
  
  while (term_count(current) > max_terms) {
    d <- drop1(current, test = "F")
    d <- d[!is.na(d$AIC), ]
    d2 <- d[rownames(d) != "<none>", ]
    
    best_term <- rownames(d2)[which.min(d2$AIC)]
    
    cat("Dropping:", best_term, 
        "| new term count =", term_count(current) - 1, "\n")
    
    current <- update(current, paste(". ~ . -", best_term))
  }
  
  current
}

aic_20_model <- trim_to_20_aic(step_aic, max_terms = 20)
formula(aic_20_model)
```

```{r}
summary(aic_20_model)
```

Final Model: aic_20_model


```{r}
par(mfrow = c(2, 2))
plot(aic_20_model)
```

## Final Model Prediction

```{r}
# Load test data
test_df <- read.csv("HousePriceTestNoY.csv")
clean_test_df <- clean_house_data(test_df)

categorical_vars <- names(which(sapply(df_clean, is.factor)))
for (col in categorical_vars) {
  if (col %in% names(clean_test_df)) {
    levels(clean_test_df[[col]]) <- levels(df_clean[[col]])
  }
}

# Predict (Log scale)
log_preds <- predict(aic_20_model, newdata = clean_test_df)

# Transform back to original scale (Dollars)
final_predictions <- exp(log_preds)

# Create submission with correct ID (Likely 'X')
submission <- data.frame(Id = clean_test_df$X, SalePrice = final_predictions)
write.csv(submission, "submission.csv", row.names = FALSE)
```
